# Reproducible Research: Peer Assessment 1

### Load required packages
```{r packages, warning=FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(ggvis)
library(ggplot2)
```

## Loading and preprocessing the data
```{r loand-prepocess}
if(!file.exists("activity.csv"))
  unzip("activity.zip")

activityDF <- read.csv("activity.csv") %>%
  mutate(date = as.Date(date, "%Y-%m-%d"))
str(activityDF)
```


## What is mean total number of steps taken per day? {#actPerDay}
```{r histogram-steps}
actPerDay <- activityDF %>%
  group_by(date) %>%
  summarise(steps = sum(steps, na.rm = TRUE))

actPerDay %>%
  ggvis(~steps/1e3) %>% 
  layer_histograms(binwidth = 1) %>%
  add_guide_axis("x", title = "Steps (thousands)") %>%
  add_guide_axis("y", title = "Day Occurence")

summarise(actPerDay, mean(steps), median(steps))
```


## What is the average daily activity pattern?
```{r average-daily-pattern}
actPerInt <- activityDF %>%
  group_by(interval) %>%
  summarise(steps = mean(steps, na.rm = TRUE))

actPerInt %>%
  ggvis(~interval, ~steps) %>%
  layer_paths() %>%
  add_guide_axis("x", title = "Interval") %>%
  add_guide_axis("y", title = "Number of Average Steps")

actPerInt$interval[ which.max(actPerInt$steps) ]
```

## Imputing missing values
```{r na-summary}
sum(is.na(activityDF$steps))
```

We are going to impute these missing values using average steps by interval.

```{r imputing-na}
imputedDF <- activityDF %>%
  left_join(actPerInt, by = "interval") %>%
  mutate(steps = ifelse(is.na(steps.x), steps.y, steps.x))
```

```{r imputed-hist}
imputedActByDay <- imputedDF %>%
  group_by(date) %>%
  summarise(steps = sum(steps))

imputedActByDay %>%
  ggvis(~steps/1e3) %>% 
  layer_histograms(binwidth = 1) %>%
  add_guide_axis("x", title = "Steps (thousands)") %>%
  add_guide_axis("y", title = "Day Occurence")

summarise(imputedActByDay, mean(steps), median(steps))
```

The histogram overall is very similar to the one we generated [earlier](#actPerDay), execept that we have far less days falling into the zero steps bucket. This makes sense because we imputed `NA` with interval average, which shifted mean up quite a bit while kept median roughly the same.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday-vs-weekend}
partitionedDF <- imputedDF %>%
  mutate(weekday = as.POSIXlt(date)$wday) %>%
  mutate(weekday = weekday > 0 & weekday < 6) %>%
  mutate(weekday = factor(weekday, labels = c("Weekend", "Weekday"))) %>%
  group_by(weekday, interval) %>%
  summarise(steps = mean(steps))

###ggvis doesn't have subplot just yet :(
ggplot(partitionedDF, aes(interval, steps)) +
  geom_line() +
  facet_grid(weekday ~ .) +
  theme_bw() +
  labs(x = "Interval",
       y = "Number of Average Steps")
```